\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-28 tex/paper.tex
save

By contrast, when running Algorithm~\ref{alg:adversarial-training} initialized
from the classifier obtained previously, in combination with a mixture density
network $r$ (modeling $Y|f(X;\theta_f)$ as a truncated mixture of four gaussian
components), adversarial training effectively reshapes the decision function
$f$, so it that becomes almost independent on $Z$, as shown in the lower plots
of Figure~\ref{fig:toy}. In particular, the conditional probability densities of
the decision scores $f(X;\theta_f)$ are now very similar to each other,
indicating only a small  dependency on the nuisance, as theoretically expected.
The dynamics of adversarial training is illustrated in
Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
making the classifier less optimal, hence the corresponding increase of ${\cal
L}_f$, but which results in a classifier that is more pivotal, hence the
corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
goes, minimizing $E$ requires making predictions that are more accurate, hence
decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
$p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
start to slighlty decrease, while remaining lower bounded by $\min_{\theta_f}
{\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
\log(2)$ of the uniform distribution over ${\cal Z}$, as shown by the dashed
line in the second plot. Finally, let us note that the ideal situation of a
classifier that is both optimal and pivotal appears to be unreachable for this
problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H(p_{Y|X}) -
\lambda H(p_Z)]$.

\glnote{Discuss practical considerations regarding the stability of training. A
better alternative seems to update $r$ until it is good
enough, then update $f$, and update $r$ again as soon as it becomes too bad.
This is important e.g. for MDN to prevent weights to explode if the network is
trained too much, and in which case gradients become useless.} \glnote{Warn against
setting $\lambda$ too high, since it could simply force $f$ to produce unlikely
and very bad decisions.}

>>>>

By contrast, when running Algorithm~\ref{alg:adversarial-training} as initialized
with the classifier obtained previously, and in combination with a mixture density
network $r$ (modeling $Z|f(X;\theta_f)$ as a truncated mixture of four gaussian
components), adversarial training effectively reshapes the decision function
$f$ so it that becomes almost independent on $Z$, as shown in the lower plots
of Figure~\ref{fig:toy}. In particular, the conditional probability densities of
the decision scores $f(X;\theta_f)$ are now very similar to each other,
indicating only a small  dependency on the nuisance, as theoretically expected.
The dynamics of adversarial training is illustrated in
Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
making the classifier less optimal, hence the corresponding increase of ${\cal
L}_f$, but which results in a classifier that is more pivotal, hence the
corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
goes, minimizing $E$ requires making predictions that are more accurate, hence
decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
$p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
start to slighlty decrease, while remaining lower bounded by $\min_{\theta_f}
{\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
\log(2)$ of the uniform distribution over ${\cal Z}$, as shown by the dashed
line in the second plot. Finally, let us note that the ideal situation of a
classifier that is both optimal and pivotal appears to be unreachable for this
problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H(p_{Y|X}) -
\lambda H(p_Z)]$.

\glnote{Discuss practical considerations regarding the stability of training.}
\glnote{Warn against
setting $\lambda$ too high, since it could simply force $f$ to produce unlikely
and very bad decisions.}
